Supplementary material to `Adaptive group-regularized logistic elastic net regression'

### 2019-07-29 docs/supplement.Rnw
just before resubmit

  \section{Introduction}\label{sec:introduction}
  This document contains supplementary material to the paper `Adaptive 
  group-regularized logistic elastic net regression'. Section 
  \ref{sec:enetprior} contains extra details on the elastic net prior 
  distribution. Section \ref{sec:objectivefunction} gives the objective 
  function that is maximised in the variational Bayes procedure introduced in
  Section \ref{md-sec:estimation} of the Main Document (MD). Section
  \ref{sec:polyagamma} gives details on
  the P{\'o}lya-Gamma parametrisation introduced in Section 
  \ref{md-sec:estimation} of the MD. The
  derivations of the variational posterior used in the MD are shown in Section
  \ref{sec:variationaldistributions}. Section
  \ref{sec:computationalcomplexity} shows how we may reduce the computational
  complexity of our variational Bayes procedure. Section 
  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
  joint log likelihood that is maximised to estimate the penalty multipliers as
  explained in Section \ref{md-sec:estimation} of the MD. The full procedure of
  the
  proposed method is shown in Section \ref{sec:fullestimationprocedure}. 
  Extensions of our method are
  presented in Section \ref{sec:extensions}. The simulations in MD Section
  \ref{md-sec:simulations} are elaborated in Section \ref{sec:simulations}. 
  Additional results to the applications in the MD Sections 
  \ref{md-sec:colorectal}--\ref{md-sec:oral} are presented in Sections 
  \ref{sec:colorectal}--\ref{sec:oral}. We
  apply our method to a study on metabolomics in Alzheimer's in Section
  \ref{sec:alzheimer} and to study in microRNAs in cervical cancer data 
  in Section \ref{sec:cervical}. Finally, 
  we introduce the ridge and lasso versions of the proposed method in Section
  \ref{sec:evbridgelasso}.

@@

  Inspection of the updating equations (\ref{md-eq:VBupdateequations}) in the 
  MD, learns us that the 
  last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
  we are left with just (\ref{eq:elboomega4}). After a change of variables 
  $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
  \begin{subequations} \label{eq:elbobeta2}
	  \begin{align}
	    \E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G 
	    |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - 
	    \frac{p}{2} \log (2 \pi) \\
	    & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in 
	    \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) 
	    (\bmu_j^2 + \bSigma_{jj}) \\
	    & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} 
	    \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	    \E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - 
	    \frac{p}{2} \log |\bSigma| .
	  \end{align}
  \end{subequations}
  We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for 
  (\ref{eq:elbobeta}):
  \begin{subequations} \label{eq:elbobeta3}
	   \begin{align}
	     \E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} 
	     [\log q (\bbeta )] \\ 
	     & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + 
	     \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + 
	     \frac{p}{2} \\
	     & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g 
	     \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 
	     \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	     & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + 
	     \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	  \end{align}
  \end{subequations}

@@

  \section{Computational complexity}\label{sec:computationalcomplexity}
  Calculation of $\bSigma$ involves the inversion of a 
  $(p \times p)$-dimensional matrix, which occurs at every iteration of the
  algorithm. If naively done, every inversion is of computational complexity
  $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck
  of the algorithm, especially with possibly many iterations. However, 
  inspection of the updating equations in (\ref{md-eq:VBupdateequations})
  in the MD, reveals
  that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
  \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the
  $c_i$ and $\chi_j$. In Section \ref{md-sec:estimation} of the MD we show that 
  for the 
  EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, 
  and the diagonal of $\bSigma$. Here we provide details on how to compute these
  with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the
  computational cost of the algorithm. If necessary, we may calculate the full
  $\bSigma$ once after the algorithm has converged.

@@

  \section{Variational Bayes expected joint log likelihood}
  \label{sec:expectedloglikelihood}
  Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then
  the variational Bayes expected joint log likelihood is derived as follows:
  \begin{align*}
    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
    \beta_j^2 | \blambda'^{(k)} \) \\ 
    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
  \end{align*}
  with the coefficients $d_g^{(k)}$ calculated as:
  \begin{align*}
    d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1}
    \beta_j^2 | \blambda'^{(k)} \) \\
    & = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} 
    ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} 
    ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} 
    ( \beta_j | \blambda'^{(k)} ) \] \\
    & = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + 
    (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} 
    \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
  \end{align*}

  \section{Full estimation procedure}\label{sec:fullestimationprocedure}
  The expectation and maximisation steps introduced in the Main Document Section
  \ref{md-sec:estimation} result
  in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
  \begin{algorithm}
    \caption{Group-regularized empirical Bayes elastic net}
    \label{alg:summaryalg}
	  \begin{algorithmic}[h]
		  \Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
		  \State{Estimate $\lambda$ by cross-validation of the regular elastic net
		  model}
		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
		  \epsilon_1$}
		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
		  for $j=1, \dots, p$ using (\ref{md-eq:VBupdateequations}) in the MD}
		  \EndWhile
		  \State{Update $\blambda'$ by (\ref{md-eq:mmlupdateequation}) in the MD}
		  \EndWhile
	  \end{algorithmic}
  \end{algorithm}

@@

  \section{Simulations}\label{sec:simulations}
  \subsection{Scenario 1}
	In this sparse scenario we assume differential signal between 5 groups of
	features. The average signal in a group increases by a factor 1.6 with group
	number:
	\begin{align*}
	  \bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} 
	  \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} 
	  & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & 
	  \beta_{50}^1 \end{matrix}}_{50}} & \cdots & 
	  \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0}
	  \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} 
	  \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr,
	  \\
	  & \, \\
	  \beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots,
	  5  \text{, } j=1, \dots, 50,
	\end{align*}
	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. We set
	$b_1 \approx 0.18$ such that the mean expected signal 
	$\overline{\E(\beta_j^g)}=0.07$. Furthermore, we
	introduce correlation between the features by setting the covariances
	$\bSigma_{ij} = 0.5^{|i-j|}$.The
	model parameters clearly do not follow the relatively heavy-tailed elastic net
	distribution. Because of this mismatch between the assumed and true 
	distributions, we expect poor performance by \texttt{gren}.

@@

  \section{Application to RNAseq in oral cancer}
  \label{sec:oral}
  In addition to the results presented in Section \ref{md-sec:oral} of the 
  MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
  \cite{simon_sparse-group_2013} and the group exponential lasso (\texttt{gel}) 
  by \cite{breheny_group_2015}. The original group lasso methods are not 
  designed to deal with overlapping groups, so in addition to using the versions
  that allow for overlapping groups, we 
  cross-tabulated the two co-data partitions to create one paritioning with 25 
  groups for \texttt{cMCP} and \texttt{gel}.
  We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
  methods that require this extra tuning parameter. The AUCs and Brier skill 
  scores calculated on the
  validation set are presented in Figures 
  \ref{fig:lines_rnaseq_oral_cancer_res1_auc}--
  \ref{fig:lines_rnaseq_oral_cancer_res1_briers}
<<lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated AUC for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
@

<<lines_rnaseq_oral_cancer_res1_briers, fig.cap="Estimated Briers skill score for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
@

@@

>>>>

  \section{Introduction}\label{sec:introduction}
  This document contains supplementary material to the paper `Adaptive 
  group-regularized logistic elastic net regression'. Section 
  \ref{sec:enetprior} contains extra details on the elastic net prior 
  distribution. Section \ref{sec:objectivefunction} gives the objective 
  function that is maximised in the variational Bayes procedure introduced in
  Section \ref{md-sec:estimation} of the Main Document (MD). Section
  \ref{sec:polyagamma} gives details on
  the P{\'o}lya-Gamma parametrisation introduced in Section 
  \ref{md-sec:estimation} of the MD. The
  derivations of the variational posterior used in the MD are shown in Section
  \ref{sec:variationaldistributions}. In Section \ref{sec:updatingequations}
  the variational and empirical Bayes updating equations for the algorithm are
  given. Section
  \ref{sec:computationalcomplexity} shows how we may reduce the computational
  complexity of our variational Bayes procedure. Extensions of our
  method are
  presented in Section \ref{sec:extensions}. The simulations in MD Section
  \ref{md-sec:simulations} are elaborated in Section \ref{sec:simulations}. 
  Additional results to the applications in the MD Sections 
  \ref{md-sec:colorectal}--\ref{md-sec:oral} are presented in Sections 
  \ref{sec:colorectal}--\ref{sec:oral}. We
  apply our method to a study on metabolomics in Alzheimer's in Section
  \ref{sec:alzheimer} and to study in microRNAs in cervical cancer data 
  in Section \ref{sec:cervical}. The computational times
  of the method are explored in Section \ref{sec:timings}. Finally, 
  we introduce the ridge and lasso versions of the proposed method in Section
  \ref{sec:evbridgelasso}.

@@

  Inspection of the updating equations (\ref{eq:VBupdateequations}), 
  learns us that the 
  last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
  we are left with just (\ref{eq:elboomega4}). After a change of variables 
  $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
  \begin{subequations} \label{eq:elbobeta2}
	  \begin{align}
	    \E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G 
	    |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - 
	    \frac{p}{2} \log (2 \pi) \\
	    & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in 
	    \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) 
	    (\bmu_j^2 + \bSigma_{jj}) \\
	    & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} 
	    \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	    \E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - 
	    \frac{p}{2} \log |\bSigma| .
	  \end{align}
  \end{subequations}
  We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for 
  (\ref{eq:elbobeta}):
  \begin{subequations} \label{eq:elbobeta3}
	   \begin{align}
	     \E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} 
	     [\log q (\bbeta )] \\ 
	     & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + 
	     \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + 
	     \frac{p}{2} \\
	     & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g 
	     \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 
	     \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	     & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + 
	     \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	  \end{align}
  \end{subequations}

@@

  \section{Updating equations}\label{sec:updatingequations}
  If we fill in the expectations and variances in the variational posterior
  from Section \ref{sec:variationaldistributions} and maximise with respect
  to the variational parameters, we arrive at the following updating equations:
  \begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	which we iterate until convergence. Here, $\bLambda'$ is a diagonal matrix 
	with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively.

	To obtain the empirical Bayes updating equations, we first derive the 
	expected log likelihood under the variational posterior as:
  \begin{align*}
    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
    \beta_j^2 | \blambda'^{(k)} \) \\ 
    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
  \end{align*}
  with the coefficients $d_g^{(k)}$ calculated as:
  \begin{align*}
    d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1}
    \beta_j^2 | \blambda'^{(k)} \) \\
    & = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} 
    ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} 
    ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} 
    ( \beta_j | \blambda'^{(k)} ) \] \\
    & = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + 
    (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} 
    \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
  \end{align*}
  Here, $\propto$ denotes proportionality with respect to $\blambda'$. The 
  resulting updating equations for the hyperparameters are now:
  \begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}

  \section{Computational complexity}\label{sec:computationalcomplexity}
  Calculation of $\bSigma$ involves the inversion of a 
  $(p \times p)$-dimensional matrix, which occurs at every iteration of the
  algorithm. If naively done, every inversion is of computational complexity
  $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck
  of the algorithm, especially with possibly many iterations. However, 
  inspection of the updating equations in (\ref{eq:VBupdateequations}), reveals
  that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
  \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the
  $c_i$ and $\chi_j$. In Section \ref{md-sec:estimation} of the MD we show that 
  for the 
  EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, 
  and the diagonal of $\bSigma$. Here we provide details on how to compute these
  with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the
  computational cost of the algorithm. If necessary, we may calculate the full
  $\bSigma$ once after the algorithm has converged.

@@

@@

  \section{Simulations}\label{sec:simulations}
  We simulated data according to five different scenarios that are explained
  in more detail below. For the regular elastic net, \texttt{gren},
  \texttt{cMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
	determines the ratio of group-level penalization ($\alpha=0$) to 
	feature level ($\alpha=1$) penalization. For the methods that do 
	automatic feature selection, we estimate models for a range of model sizes:
	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
	from $p=900$ to $p=1000$. The covariance matrix of the features and 
	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
  Scenarios. We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report average area under the receiver operator curve (AUC) and 
	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
	measures of discrimination and calibration, respectively, and for both a 
	higher value indicates better predictive performance.
	For the methods with an $\alpha$ parameter, we display the best-performing 
	$\alpha$. Results for the other $\alpha$'s are presented in 
	Figures \ref{fig:lines2_simulations_res1_auc_briers}-
	\ref{fig:lines2_simulations_res5_auc_briers}-

<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (i).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (ii).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (iii).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (iv).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (v).", out.width="100%", fig.asp=1/2>>=
@

  \subsection{Scenario 1}
	In this sparse scenario we assume differential signal between 5 groups of
	features. The average signal in a group increases by a factor 1.6 with group
	number:
	\begin{align*}
	  \bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} 
	  \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} 
	  & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & 
	  \beta_{50}^1 \end{matrix}}_{50}} & \cdots & 
	  \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0}
	  \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} 
	  \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr,
	  \\
	  & \, \\
	  \beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots,
	  5  \text{, } j=1, \dots, 50,
	\end{align*}
	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. We set
	$b_1 \approx 0.18$ such that the mean expected signal 
	$\overline{\E(\beta_j^g)}=0.07$. Furthermore, we
	introduce correlation between the features by setting the covariances
	$\bSigma_{ij} = 0.5^{|i-j|}$.The
	model parameters clearly do not follow the relatively heavy-tailed elastic net
	distribution. Because of this mismatch between the assumed and true 
	distributions, we expect poor performance by \texttt{gren}.

@@

  \section{Application to RNAseq in oral cancer}
  \label{sec:oral}
  In addition to the results presented in Section \ref{md-sec:oral} of the 
  MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
  \cite{simon_sparse-group_2013} and the group exponential lasso (\texttt{gel}) 
  by \cite{breheny_group_2015}. The original group lasso methods are not 
  designed to deal with overlapping groups, so in addition to using the versions
  that allow for overlapping groups, we 
  cross-tabulated the two co-data partitions to create one paritioning with 25 
  groups for \texttt{cMCP} and \texttt{gel}.
  We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
  methods that require this extra tuning parameter. The AUCs and Brier skill 
  scores calculated on the
  validation set are presented in Figures 
  \ref{fig:lines_rnaseq_oral_cancer_metastasis_res1_auc}--
  \ref{fig:lines_rnaseq_oral_cancer_metastasis_res1_briers}
<<lines_rnaseq_oral_cancer_metastasis_res1_auc, fig.cap="Estimated AUC for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
@

<<lines_rnaseq_oral_cancer_metastasis_res1_briers, fig.cap="Estimated Briers skill score for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
@

@@

  \section{Computational times}\label{sec:timings}
<<>>=
bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
                     stringsAsFactors=FALSE)

bench2 <- read.table("results/rnaseq_oral_cancer_metastasis_bench1.csv", 
                     stringsAsFactors=FALSE)
times1 <- round(bench1$time*10^(-9), 2)
times2 <- round(bench2$time*10^(-9), 2)
@
  On the colorectal cancer example in MD Section \ref{md-sec:colorectal}
  we achieved the following times:
  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively. \texttt{GRridge} and \texttt{SGL} with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
  respectively.
  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
  The results for the oral cancer example in MD Section \ref{md-sec:oral} 
  are similar. \texttt{gren} 
  achieved 
  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively.
  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}.
